---
title: "SwarupWang_ForecastCompetitionReport"
subtitle: "[GitHub](https://github.com/popping-rosemary/SwarupWang_ENV797_TSA_ForecastCompetition_S25.git)"
author: "Lucy Wang, Sameer Swarup"
date: "2025-04-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Load Packages 
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(smooth)
library(zoo)
library(kableExtra)
library(knitr)
```

## Data Wrangling

### Load load, temp, and humidity files 
```{r data file}
# Load the demand excel file
load_df <- read_excel("./Data/load.xlsx")
  
daily_load <- load_df %>% 
  group_by(meter_id, date) %>%
  summarize(
    daily_avg = mean(c_across(1:24), na.rm = TRUE),
    .groups   = "drop"
  )

# Load the temperature file
temp_df <- read_excel("./Data/temperature.xlsx")

daily_temp <- temp_df[1:3] %>% 
  group_by(date) %>%
  summarize(
    daily_avg = mean(t_ws1, na.rm = TRUE),
    .groups   = "drop"
  )

# Load the humidity file
humidity_df <- read_excel("./Data/relative_humidity.xlsx")

daily_humidity <- humidity_df[1:3] %>% 
  group_by(date) %>%
  summarize(
    daily_avg = mean(rh_ws1, na.rm = TRUE),
    .groups   = "drop"
  )
```

## load file for humidity and temperature forecasts

The forecasting code can be found in the Code folder 

```{r load file for humidity and temperature forecasts}
daily_temp_for <- read.csv("./Data/Temp_Forecast.csv") 
temp_for <- msts(daily_temp_for$x, seasonal.periods =c(7,365.25), start=c(2011,1,1))

daily_humidity_for <- read.csv("./Data/Humidity_Forecast.csv") 
humidity_for <- msts(daily_humidity_for$x, seasonal.periods =c(7,365.25), start=c(2011,1,1))
```

## Create Time Series

```{r create TS}
# Create time series for load 
load_ts <- msts(daily_load$daily_avg, 
                seasonal.periods =c(7,365.25), 
                start=c(2005,1,1))

# Create time series for temperature 
temp_ts <- msts(daily_temp$daily_avg, 
                seasonal.periods =c(7,365.25), 
                start=c(2005,1,1))

# Create time series for humidity
humidity_ts <- msts(daily_humidity$daily_avg, 
                seasonal.periods =c(7,365.25), 
                start=c(2005,1,1))
```

```{r}
#create a subset for training purpose
n_for = 365
ts_load_train <- subset(load_ts,
                        end = length(load_ts)-n_for)

#create a subset for testing purpose
ts_load_test <- subset(load_ts,
                       start = length(load_ts)-n_for)

# Create date sequence
dates <- seq.Date(from = as.Date("2011-01-01"), to = as.Date("2011-02-28"), by = "day")

```

## Best Five Models 

### Model 1: Forecasting using NN - p=1, P=0, K=c(2,4)

```{r model 1}
K = c(2,4)
for_period = 59

NN_forecast <- function(ts, p, P, K){
  model_fit <- nnetar(ts,
                 p=p,
                 P=P,
                 xreg=fourier(ts, K=K))
  
  model_forcast <-  forecast(model_fit, h=59,xreg=fourier(ts, 
                                          K=K,h=for_period))
}

forecast_1 <- NN_forecast(load_ts, 1, 0, K)

#Plot model + observed data
autoplot(load_ts, color="gray70") +
  autolayer(forecast_1, series="Neural Network",PI=FALSE)+
  ylab("Load") +
  ggtitle("Daily Load Forecast - Model 1")+
  theme_minimal()+
  theme(legend.position = "bottom")

NN_result_1 <- as.data.frame(forecast_1$mean)

# Combine into a dataframe
print_result <- function(result){
  df <- data.frame(Date = dates, Value = result) # Combine into a dataframe
  df_summary <- bind_rows(
  head(df),
  tail(df)
) # Get head and tail
  # Print a table
  kable(df_summary, caption = "Model result for the first and last five days") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
}

print_result(NN_result_1)
```

### Model 2: Forecasting using NN - p=1, P=1, K=c(2,4)

```{r model 2}
forecast_2 <- NN_forecast(load_ts, 1, 1, K)

#Plot model + observed data
autoplot(load_ts, color="gray70") +
  autolayer(forecast_2, series="Neural Network",PI=FALSE)+
  ylab("Load") +
  ggtitle("Daily Load Forecast - Model 2")+
  theme_minimal()+
  theme(legend.position = "bottom")

NN_result_2 <- as.data.frame(forecast_2$mean)

print_result(NN_result_2)
```

### Model 3: Forecasting using TBATS+NN Combined Model with humidity and temperature variables

```{r model 3}
h_future <- 59
K=c(2,4)
# Generate Fourier terms for training (if needed) and for forecasting 2010
fourier_train <- fourier(load_ts, K = K)
fourier_future <- fourier(load_ts, K = K, h = h_future)

# Create xreg matrices for training and forecasting
# (Assumes you have also created ts_temp_train and ts_humidity_train for training period)
xreg_train <- cbind(fourier_train, temp = temp_ts, hum = humidity_ts)
xreg_future <- cbind(fourier_future, temp = temp_for, hum = humidity_for)

# Fit the NNAR model on training data with external regressors
NN_fit <- nnetar(load_ts, p = 1, P = 1, xreg = xreg_train)

# Forecast for h_future days (2010)
NN_for <- forecast(NN_fit, xreg = xreg_future, h = h_future)

# TBATS can take time to fit
TBATS_fit <- tbats(load_ts)

TBATS_for <- forecast(TBATS_fit, h=59)

hybrid_for <- (TBATS_for$mean + NN_for$mean) / 2

autoplot(load_ts, color="gray70") +
  autolayer(hybrid_for, series="Neural Network",PI=FALSE)+
  ylab("Load") +
  ggtitle("Daily Load Forecast - Model 3")+
  theme_minimal()+
  theme(legend.position = "bottom")

hybrid_result <- as.data.frame(hybrid_for)
print_result(hybrid_result)
```

### Model 4:  Forecasting using further combined model based on Model 3
```{r model 4}
combined_for <- (TBATS_for$mean + forecast_2$mean) / 2

forecast_4 <- (combined_for + NN_for$mean) / 2

autoplot(load_ts, color="gray70") +
  autolayer(forecast_4, series="Neural Network",PI=FALSE)+
  ylab("Load") +
  ggtitle("Daily Load Forecast - Model 4")+
  theme_minimal()+
  theme(legend.position = "bottom")

model4_result <- as.data.frame(forecast_4)
print_result(model4_result)
```
### Model 5: Forecasting using NN - p=1, P=0, K=c(2,2)

```{r model 5}
K=c(2,2)
forecast_5 <- NN_forecast(load_ts, 1, 0, K)

#Plot model + observed data
autoplot(load_ts, color="gray70") +
  autolayer(forecast_2, series="Neural Network",PI=FALSE)+
  ylab("Load") +
  ggtitle("Daily Load Forecast - Model 5")+
  theme_minimal()+
  theme(legend.position = "bottom")

NN_result_5 <- as.data.frame(forecast_5$mean)

print_result(NN_result_5)
```

